The combination of reporting bias and underpowered study designs has substantially exaggerated the motor learning benefits of self-controlled practice and enhanced expectancies: A meta-analysis

### 2023-02-07 docs/revision-1/r1_manuscript.tex
init

For nearly 20 years, motor learning and performance research has been accumulating that some (e.g., Wulf et al., 2010; Wulf \& Lewthwaite, 2016, 2021) have argued is not well-explained by classic information-processing-based theories (Guadagnoli \& Lee, 2004; Salmoni et al., 1984; Schmidt, 1975). Examples range from putative motor learning benefits from receiving bogus social comparison feedback (e.g., Ávila et al., 2012) to ostensive learning and performance enhancement as a consequence of making incidental choices during practice (e.g., Lewthwaite et al., 2015). In the Optimizing Performance Through Intrinsic Motivation and Attention for Learning (OPTIMAL) theory, Wulf and Lewthwaite (2016) argued that the failure of information-processing theories to account for findings like these created the need for a new theory of motor learning. The OPTIMAL theory provided an account of these, and numerous other lines of research, proposing that autonomy support and enhanced expectancies increase motivation, which explains the learning and performance benefits observed with self-controlled practice (see Sanli et al., 2013 for a review), positive feedback (e.g., Chiviacowsky \& Wulf, 2007), social comparative feedback (e.g., Ávila et al., 2012), perceived task difficulty (e.g., C. Lee et al., 2011), conceptions of ability (e.g., Cimpian et al., 2007), self-modeling (e.g., Ste-Marie et al., 2011), and extrinsic rewards (e.g., Abe et al., 2011). While motivation had been deemphasized in previous motor learning theories, within OPTIMAL theory numerous lines of 21st century research may be best explained by motivational rather than informational mechanisms.

Unfortunately, it has become recognized throughout social science that reporting bias in the literature can seriously distort the evidence (Collaboration, 2015; Fanelli, 2010; Hagger et al., 2016; Munafò et al., 2017). The results explained by the motivational factors in OPTIMAL theory may therefore be exaggerated or even non-existent (Gelman \& Carlin, 2014). Research on the motor learning literature specifically has found evidence of the pernicious combination of low statistical power, high multiplicity (i.e., many statistical tests), and selective reporting (Lohse et al., 2016). The evidence underpinning the autonomy support predictions in OPTIMAL theory has been questioned in a recent meta-analysis, suggesting that low power combined with reporting bias may be responsible for the apparent benefits of self-controlled practice (McKay et al., in-press). In that study, it was observed that published experiments found an average self-controlled practice benefit of \emph{g} = .54. However, several models of reporting bias provided a better fit to the data than the naive random effects model. Each of the models suggested the true average effect was small or potentially zero. Self-controlled practice was the primary literature addressed by the autonomy support factor in OPTIMAL theory; yet, these findings call into question whether there was even a robust phenomenon to explain.

Bacelar and colleagues (2022) investigated the literature addressed by the second motivation factor in OPTIMAL theory---enhanced expectancies. In their meta-analysis, Bacelar and colleagues found that the average benefit of studies that manipulated expectancies via interventions described in OPTIMAL theory was \emph{g} = .54. However, there was evidence of reporting bias that could not be accounted for with moderators in the study. The authors applied the trim-and-fill method in an effort to adjust for reporting bias, but it made no corrections and no other corrections were applied. It is noteworthy that self-controlled practice studies and the studies included in the Bacelar et al.~meta-analysis have much in common: both literatures are comprised of variables predicted by OPTIMAL theory to increase motivation and in turn motor performance and learning; both include studies examined in a meta-analysis by Lohse and colleagues (2016) that found evidence of low power, multiplicity, and bias; and both showed signs of reporting bias in their funnel plots. If both literatures have been affected similarly by reporting bias, then the current estimate of \emph{g} = .54 for variables thought to enhance expectancies may be a substantial overestimate.

Addressing reporting bias presents substantial challenges to meta-analysts. Since reporting bias limits the information we have access to, it is impossible to know for certain how much bias is present or how large the impact is (Carter et al., 2019; McShane et al., 2016). The best we can do is think carefully about the mechanisms that potentially underlie reporting bias and attempt to model them accurately. In contrast, a naive random effects analysis assumes there is no reporting bias. Complicating matters, reporting bias can take several different forms with unique impacts on the final sample (Maier et al., in-press; Stefan \& Schönbrodt, 2022; Thornton \& Lee, 2000). To account for this, multiple models of reporting bias need to be attempted without knowing which is most likely \emph{a priori}. Thus, bias-correction analyses are inherently sensitivity analyses (Mathur \& VanderWeele, 2020; Sutton et al., 2000; Vevea \& Woods, 2005). Until recently, if the results of multiple sensitivity analyses differed widely, there was no mechanism to reconcile the estimates. We now have Robust Bayesian Meta-Analysis (RoBMA) methods that apply Bayesian model averaging to allow meta-analysts to fit several plausible models (see Table 1 for descriptions) and give greater weight to the models that best account for the data (Bartoš et al., 2022; Maier et al., in-press). The RoBMA method provides single estimates of the average effect and heterogeneity, along with Bayes factors to quantify the evidence in support of a true effect, the presence of heterogeneity, and the presence of reporting bias. Results from simulation studies and analyses of real data with known reporting bias mechanisms suggest that RoBMA is substantially more accurate and less biased than naive random effects models and also performs better than other competing reporting bias models (Bartoš et al., 2022).

@@

A challenge for most reporting bias models is large heterogeneity in true effects. While RoBMA appears to perform well with moderate levels of heterogeneity, its performance has not been evaluated when heterogeneity is high and the performance of each of its constituent reporting bias models suffers with high heterogeneity (Carter et al., 2019). The \emph{z}-curve model was designed specifically to perform well regardless of heterogeneity (Bartoš \& Schimmack, 2022). While \emph{z}-curve does not provide adjusted effect size estimates (and such estimates may be meaningless with high heterogeneity), it instead estimates the average underlying power of included experiments. A significant difference between the estimated power of studies and the observed proportion of significant results can indicate the presence of reporting bias in a literature.

Here, we leveraged state-of-the-art robust Bayesian meta-analysis and \emph{z}-curve methods to re-analyze the meta-analyses by McKay and colleagues (in-press) and Bacelar and colleagues (2022). Considering the potential importance of OPTIMAL theory for the field of motor learning and performance (see T. D. Lee \& Carnahan, 2021 for a discussion), it is imperative that the evidence buttressing its motivation predictions be evaluated as rigorously as possible. Critically, our re-analysis addresses limitations in both previous meta-analyses. First, the analysis of enhanced expectancies fit only one bias correction model---the trim-and-fill method (Duval \& Tweedie, 1998)---and that model has been shown to result in exaggerated effect size estimates and severely inflated Type 1 error rates in the presence of publication bias and small or null effects (Bartoš et al., 2022; Carter et al., 2019; Hong \& Reed, 2021). Second, although the results from multiple reporting bias models coalesced around small effect sizes (ranged from \emph{g} = -.11 to \emph{g} = .26) in the analysis of self-controlled practice, there are no principled reasons for preferring one estimate over another. We now consider a wider range of plausible models of reporting bias than that used in the previous meta-analyses. We also leverage Bayesian model averaging to upweight the best performing models, which has the advantage of evaluating single model-averaged posterior distributions for each parameter of interest. Lastly, we fit \emph{z}-curve models to the data in both meta-analyses. With this technique, the average power can be estimated and compared to the rate of significant results, providing crucial insight into the quality of the evidence-base supporting motivation predictions in OPTIMAL theory.

@@

The original meta-analyses followed similar inclusion criteria and data extraction policies (see Figure 1). All data were extracted by two independent researchers with conflicts settled by including a third researcher. The literature search for the self-controlled practice meta-analysis (for further details see McKay et al., in-press) ended in August 2019 and in June 2020 for the enhanced expectancies meta-analysis (for further details see Bacelar, Parma, Murrah, et al., 2022).

@@

The self-controlled practice and enhanced expectancy meta-data were reanalyzed using robust Bayesian meta-analysis with publication selection model-averaging (RoBMA-PSMA) (Maier et al., in-press). The RoBMA-PSMA method evaluates the evidence for reporting bias and adjusts effect size estimates and 95\% credible intervals to account for the estimated bias. Since the true data generating process underlying the data is unknown, the RoBMA-PSMA method fits several plausible models to the data. The models vary with respect to whether a) an effect is present or absent, b) there is one true effect (fixed effect) or a true average effect (random effects), c) reporting bias is present or absent, and d) if present, how results are selected for publishing.

@@

RoBMA includes 36 separate models: a) six weight-function selection models in each level of a 2 (Effect: present, absent) x 2 (Heterogeneity: present, absent) matrix (24 models in total), b) PET and PEESE regression models in each level of the 2 x 2 Effect x Heterogeneity matrix (8 models in total), and c) models assuming no reporting bias at each level of the Effect x Heterogeneity matrix (4 models in total). The prior probabilities for the reporting-bias adjusted and unadjusted models are both set to .5; thus, summing to 1.0. The prior for the reporting-bias adjusted models is spread evenly between the selection and PET-PEESE regression model classes, so the priors for the two model classes sum to .5. Estimates from all 36 models are combined using Bayesian model-averaging, which weights each model estimate based on its posterior probability. Models that better account for the data are given greater weight in the RoBMA-PSMA model and models that provide a poor fit are down-weighted. The adjusted effect size estimate is produced by averaging across all models. This preserves the uncertainty about the true data generating process while weighting the component models based on their relative performance.

In addition to providing an overall effect estimate, the RoBMA-PSMA model can also be used to calculate Bayes factors (BF) quantifying the strength of support for the presence (or absence) of an intervention effect, heterogeneity, or reporting bias. To illustrate, consider the question of whether there is or is not an intervention effect. Since we used a neutral prior that considered each possibility equally likely, we can take the ratio of the posterior probabilities of model ensembles that included an effect to those that did not. We always report the BFs so they can be interpreted as how many times more likely the data were assuming the best supported hypothesis compared to the other hypothesis. For example, a BF01 = 2.0 for the absence of an intervention effect suggests that the models assuming all variation is random fit the data twice as well as models predicting an effect is present\footnote{BF\textsubscript{10} is the inverse of BF\textsubscript{01}, so while BF\textsubscript{01} indicates evidence in support of the null hypothesis, BF\textsubscript{10} indicates evidence in support of the alternative hypothesis.}. Bayes factors should be interpreted as a continuous measure of the relative support for one hypothesis versus another, not as the probability a hypothesis is true.

RoBMA-PSMA outperforms other meta-analytic approaches, including each constituent model included in RoBMA-PSMA, in simulations covering a range of plausible scenarios (Maier et al., in-press). Further, an analysis of real data from an approximately known data generating process (multi-lab registered replication report) suggested that RoBMA-PSMA outperforms other available techniques (Bartoš et al., 2022). These strong performance indices combined with the ability to simultaneously model various plausible manifestations of reporting bias made RoBMA-PSMA an attractive choice for re-analyzing the enhanced expectancies and self-controlled practice meta-analyses.

@@

The self-controlled practice meta-data\footnote{\emph{Z}-curve results for self-controlled practice were reported previously in McKay et al. (in-press). We reproduce them here for comparison to enhanced expectancies and motivational factors combined analyses.}, enhanced expectancy meta-data, and the combination of both the enhanced expectancy and self-controlled practice meta-data were analyzed with a \emph{z}-curve. A \emph{z}-curve analysis estimates the statistical power of all studies ever conducted within a given literature, even if those studies were not reported, on the basis of the significant results that are present (Bartoš et al., 2022). That power estimate is equivalent to the expected discovery rate, that is, the expected rate of significant results for a given literature. The expected discovery rate estimated by a \emph{z}-curve and its corresponding 95\% confidence interval can be compared to the observed discovery rate in the literature (the actual rate of significant results). A discrepancy between the 95\% confidence interval of the expected discovery rate estimate and the observed discovery rate provides evidence of reporting bias.

@@

Similar to RoBMA-PSMA (Maier et al., in-press), \emph{z}-curve has also performed well in simulation studies and when applied to multi-lab registered replication data (Bartoš \& Schimmack, 2022). \emph{Z}-curve is a tool that can provide insight into the possible power shortcomings of a particular literature. Further, as \emph{z}-curve was designed to accommodate highly heterogeneous inputs, it is ideal for exploring power and reporting bias among studies that used a variety of methods to manipulate the two motivational factors in Wulf and Lewthwaite's (2016) OPTIMAL theory of motor learning.

@@

We fit two RoBMA-PSMA models to the enhanced expectancies meta-data using the effect sizes and variances calculated by Bacelar and colleagues (2022). The first model included all studies, while the second excluded two influential cases, consistent with the primary results reported in the original meta-analysis. The self-controlled practice effect sizes and standard errors reported by McKay and colleagues (in-press) were analyzed using the same strategy.

@@

Statistical analyses were conducted using R (Version 4.1.2; R Core Team, 2021) and the R-packages \emph{geomtextpath} (Version 0.1.0; Cameron \& van den Brand, 2022), \emph{gt} (Version 0.6.0; Iannone et al., 2022), \emph{invgamma} (Version 1.1; Kahle \& Stamey, 2017), \emph{metafor} (Version 3.4.0; Viechtbauer, 2010), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{patchwork} (Version 1.1.0.9000; Pedersen, 2022), \emph{plotly} (Version 4.10.0; Sievert, 2020), \emph{pwr} (Version 1.3.0; Champely, 2020), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{robma} (Bartoš \& Maier, 2020), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{truncdist} (Version 1.0.2; Novomestky \& Nadarajah, 2016), and \emph{zcurve} (Version 2.1.2; Bartoš \& Schimmack, 2020) were used in this project.

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure 2B. The results suggest moderate evidence against the presence of an effect, BF\textsubscript{01} = 3.16, very weak evidence against the presence of heterogeneity, BF\textsubscript{rf} = 1.7, and overwhelming evidence for the presence of reporting bias, BF\textsubscript{pb} = 18,399. The overall model ensemble estimated the effect of self-controlled practice as \emph{d} = .034 (95\% credible interval {[}.0, .248{]}). Heterogeneity was estimated as \(\tau\) = .05 (95\% credible interval {[}.0, .261{]}). A model fit with two influential cases included found overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 1,924,516 and estimated \(\tau\) = .559 (95\% credible interval {[}.36, .78{]}). There were no other meaningful differences between models.

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure 2C. The results revealed weak evidence for the presence of an effect BF\textsubscript{10} = 1.9, weak evidence for the presence of reporting bias, BF\textsubscript{pb} = 2.3, and very strong evidence for the presence of heterogeneity, BF\textsubscript{rf} = 47.6. The model ensemble estimated an average effect of \emph{d} = .26 (95\% credible interval {[}-.07, .63{]}). Heterogeneity was estimated as \(\tau\) = .35 (95\% credible interval {[}.07, .54{]}). A model fit with two influential cases included found weak evidence for the absence of an effect BF\textsubscript{01} = 2.5, strong evidence for the presence of reporting bias, BF\textsubscript{pb} = 21, and overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 45,300. The model ensemble estimated an average effect of \emph{d} = .00 (95\% credible interval {[}-.62, .59{]}), \(\tau\) = .49 (95\% credible interval {[}.32, .68{]}).

The RoBMA-PSMA models with and without influential cases differed primarily with respect to the fit of the PEESE models. When two large effect sizes with large standard errors were included in the analysis, the PEESE model provided a very strong fit to the data (BF\textsubscript{10} = 39.9) when assuming no true effect but the presence of heterogeneity. When the two influential cases were excluded, the best fitting model was the PET under the same assumptions (BF\textsubscript{10} = 9.4).

@@

\leavevmode\vadjust pre{\hypertarget{ref-R-robma}{}}%
Bartoš, F., \& Maier, M. (2020). \emph{RoBMA: An r package for robust bayesian meta-analyses}. \url{https://CRAN.R-project.org/package=RoBMA}

@@

\leavevmode\vadjust pre{\hypertarget{ref-maier2022}{}}%
Maier, M., Bartoš, F., \& Wagenmakers, E.-J. (in-press). Robust {Bayesian} meta-analysis: Addressing publication bias with model-averaging. \emph{Psychological Methods}. \url{https://doi.org/10.1037/met0000405}

>>>>

For nearly 20 years, motor learning and performance research has been accumulating that some (e.g., Wulf et al., 2010; Wulf \& Lewthwaite, 2016, 2021) have argued is not well-explained by classic information-processing-based theories (Guadagnoli \& Lee, 2004; Salmoni et al., 1984; Schmidt, 1975). Examples range from putative motor learning benefits from receiving bogus social comparison feedback (e.g., Ávila et al., 2012) to ostensive learning and performance enhancement as a consequence of making incidental choices during practice (e.g., Lewthwaite et al., 2015). In the Optimizing Performance Through Intrinsic Motivation and Attention for Learning (OPTIMAL) theory, Wulf and Lewthwaite (2016) argued that \textcolor{blue}{information processing theories account for only a temporary influence of motivation, but fail to capture the more permanent influences of motivation on motor learning suggested by these findings; this failure to account for more permanent motor learning effects created the need for a new theory.} The OPTIMAL theory provided an account of these, and numerous other lines of research, proposing that autonomy support and enhanced expectancies increase motivation, which explains the learning and performance benefits observed with self-controlled practice (see Sanli et al., 2013 for a review), positive feedback (e.g., Chiviacowsky \& Wulf, 2007), social comparative feedback (e.g., Ávila et al., 2012), perceived task difficulty (e.g., C. Lee et al., 2011), conceptions of ability (e.g., Cimpian et al., 2007), self-modeling (e.g., Ste-Marie et al., 2011), and extrinsic rewards (e.g., Abe et al., 2011). While motivation had been deemphasized in previous motor learning theories, within OPTIMAL theory numerous lines of 21st century research may \textcolor{blue}{also be explained by motivational mechanisms, rather than or in addition to informational mechanisms.}

Unfortunately, it has become recognized throughout social science that reporting bias in the literature can seriously distort the evidence (Collaboration, 2015; Fanelli, 2010; Hagger et al., 2016; Munafò et al., 2017). The results explained by the motivational factors in OPTIMAL theory may therefore be exaggerated or even non-existent (Gelman \& Carlin, 2014). Research on the motor learning literature specifically has found evidence of the pernicious combination of low statistical power, high multiplicity (i.e., many statistical tests), and selective reporting (Lohse et al., 2016). \textcolor{blue}{Therefore, it is necessary to carefully examine the literature supporting OPTIMAL theory predictions to identify and potentially correct for selective reporting and publication.}

The evidence underpinning the autonomy support predictions in OPTIMAL theory has been questioned in a recent meta-analysis, suggesting that low power combined with reporting bias may be responsible for the apparent benefits of self-controlled practice (McKay, Yantha, et al., 2022). In that study, it was observed that published experiments found an average self-controlled practice benefit of \emph{g} = .54. However, several models of reporting bias provided a better fit to the data than the naive random effects model. Each of the models suggested the true average effect was small or potentially zero. Self-controlled practice was the primary literature addressed by the autonomy support factor in OPTIMAL theory; yet, these findings call into question whether there was even a robust phenomenon to explain.

Bacelar, Parma, Murrah, et al. (2022) investigated the literature addressed by the second motivation factor in OPTIMAL theory---enhanced expectancies. In their meta-analysis, it was found that the average benefit of studies that manipulated expectancies via interventions described in OPTIMAL theory was \textcolor{blue}{also} \emph{g} = .54. However, there was evidence of reporting bias that could not be accounted for with moderators in the study. The authors applied the trim-and-fill method in an effort to adjust for reporting bias, but it made no corrections and no other corrections were applied. It is noteworthy that self-controlled practice studies and the studies included in the \textcolor{blue}{enhanced expectancies} meta-analysis have much in common \textcolor{blue}{(in addition to identical random effects point estimates based on the published literature)}: a) both literatures are comprised of variables predicted by OPTIMAL theory to increase motivation and in turn motor performance and learning; b) both include \textcolor{blue}{a study or two} examined in a meta-analysis by Lohse and colleagues (2016) that found evidence of low power, multiplicity, and bias; and c) both showed signs of \textcolor{blue}{small study effects} in their funnel plots. If both literatures have been affected similarly by reporting bias, then the current estimate of \emph{g} = .54 for variables thought to enhance expectancies may be a substantial overestimate.

Addressing reporting bias presents substantial challenges to meta-analysts. Since reporting bias limits the information we have access to, it is impossible to know for certain how much bias is present or how large the impact is (Carter et al., 2019; McShane et al., 2016). The best we can do is think carefully about the mechanisms that potentially underlie reporting bias and attempt to model them accurately. In contrast, a naive random effects analysis assumes there is no reporting bias. Complicating matters, reporting bias can take several different forms with unique impacts on the final sample (Maier et al., 2022; Stefan \& Schönbrodt, 2022; Thornton \& Lee, 2000). To account for this, multiple models of reporting bias need to be attempted without knowing which is most likely \emph{a priori}. Thus, bias-correction analyses are inherently sensitivity analyses (Mathur \& VanderWeele, 2020; Sutton et al., 2000; Vevea \& Woods, 2005). Until recently, if the results of multiple sensitivity analyses differed widely, there was no mechanism to reconcile the estimates. We now have Robust Bayesian Meta-Analysis (RoBMA) methods that apply Bayesian model averaging to allow meta-analysts to fit several plausible models (see Table 1 for descriptions) and give greater weight to the models that best account for the data (Bartoš et al., 2022; Maier et al., 2022). The RoBMA method provides single estimates of the average effect and heterogeneity, along with Bayes factors to quantify the evidence in support of a true effect, the presence of heterogeneity, and the presence of reporting bias. Results from simulation studies and analyses of real data with known reporting bias mechanisms suggest that RoBMA is substantially more accurate and less biased than naive random effects models, and also performs better than other competing reporting bias models (Bartoš et al., 2022).

@@

A challenge for most reporting bias models is large heterogeneity in true effects (Carter et al., 2019). While RoBMA appears to perform well with moderate levels of heterogeneity, its performance has not been evaluated when heterogeneity is high and the performance of each of its constituent reporting bias models suffers with high heterogeneity. The \emph{z}-curve model was designed specifically to perform well regardless of heterogeneity (Bartoš \& Schimmack, 2022). While \emph{z}-curve does not provide adjusted effect size estimates (and such estimates may be meaningless with high heterogeneity), it instead estimates the average underlying power of included experiments. A significant difference between the estimated power of studies and the observed proportion of significant results can indicate the presence of reporting bias in a literature.

Here, we leveraged state-of-the-art robust Bayesian meta-analysis and \emph{z}-curve methods to re-analyze the meta-analyses by McKay and colleagues (2022) and Bacelar and colleagues (2022). Considering the potential importance of OPTIMAL theory for the field of motor learning and performance (see T. D. Lee \& Carnahan, 2021 for a discussion), it is imperative that the evidence buttressing its motivation predictions be evaluated as rigorously as possible. Critically, our re-analysis addresses limitations in both previous meta-analyses. First, the analysis of enhanced expectancies fit only one bias correction model---the trim-and-fill method (Duval \& Tweedie, 1998)---and that model has been shown to result in exaggerated effect size estimates and severely inflated Type 1 error rates in the presence of publication bias and small or null effects (Bartoš et al., 2022; Carter et al., 2019; Hong \& Reed, 2021). Second, although the results from multiple reporting bias models coalesced around small effect sizes (ranged from \emph{g} = -.11 to \emph{g} = .26) in the analysis of self-controlled practice, there are no principled reasons for preferring one estimate over another. We now consider a wider range of plausible models of reporting bias than that used in the previous meta-analyses. We also leverage Bayesian model averaging to upweight the best performing models, which has the advantage of evaluating single model-averaged posterior distributions for each parameter of interest. Lastly, we fit \emph{z}-curve models to the data in both meta-analyses. With this technique, the average power can be estimated and compared to the rate of significant results, providing crucial insight into the quality of the evidence-base supporting motivation predictions in OPTIMAL theory.

@@

The original meta-analyses followed similar inclusion criteria and data extraction policies (see Figure 1). All data were extracted by two independent researchers with conflicts settled by including a third researcher. The literature search for the self-controlled practice meta-analysis (for further details see McKay, Yantha, et al., 2022) ended in August 2019 and in June 2020 for the enhanced expectancies meta-analysis (for further details see Bacelar, Parma, Murrah, et al., 2022).

@@

The self-controlled practice and enhanced expectancy meta-data were reanalyzed using robust Bayesian meta-analysis with publication selection model-averaging (RoBMA-PSMA) (Bartoš et al., 2022). The RoBMA-PSMA method evaluates the evidence for reporting bias and adjusts effect size estimates and 95\% credible intervals to account for the estimated bias. Since the true data generating process underlying the data is unknown, the RoBMA-PSMA method fits several plausible models to the data. The models vary with respect to whether a) an effect is present or absent, b) \textcolor{blue}{homogeneous studies are summarized with a common effect (fixed effects) or a distribution of heterogeneous studies summarized with a mean effect and heterogeneity (random effects)}, c) reporting bias is present or absent, and d) if present, how results are selected for publishing.

@@

RoBMA includes 36 separate models: a) six weight-function selection models in each level of a 2 (Effect: present, absent) x 2 (Heterogeneity: present, absent) matrix (24 models in total), b) PET and PEESE regression models in each level of the 2 x 2 Effect x Heterogeneity matrix (8 models in total), and c) models assuming no reporting bias at each level of the Effect x Heterogeneity matrix (4 models in total). The prior probabilities for the reporting-bias adjusted and unadjusted models are both set to .5; thus, summing to 1.0. The prior for the reporting-bias adjusted models is spread evenly between the selection and PET-PEESE regression model classes, so the priors for the two model classes sum to .5. \textcolor{blue}{Similarly, the prior distribution for the presence of an effect was divided evenly between a spike prior equal to zero effect (i.e., the null) and a normal distribution of plausible true effects with $M$ = 0 and $SD$ = 1. Likewise, the prior probabilities of a fixed effect and heterogenous effects were divided evenly, with plausible values of $\tau$ following an inverse gamma distribution with shape = 1 and scale = .15. We chose neutral priors}\footnote{\textcolor{blue}{We considered a more informed distribution of plausible effects based on the empirical benchmarks reported by Lovakov and Agadullina} (2021) \textcolor{blue}{for social psychological phenomena. In this sensitivity analysis, we included a normal prior distribution with $M$ = 0 and $SD$ = .36. The results of this sensitivity analysis were nearly identical to the primary analysis.}} \textcolor{blue}{to allow the data to dominate our posterior estimates and not impose our own opinions on the analysis. Admittedly, our prior beliefs were that publication bias is likely, so the analysis we report is somewhat conservative in this respect relative to our personal priors.} Estimates from all 36 models are combined using Bayesian model-averaging, which weights each model estimate based on its posterior probability. Models that better account for the data are given greater weight in the RoBMA-PSMA model and models that provide a poor fit are down-weighted. The adjusted effect size estimate is produced by averaging across all models. This preserves the uncertainty about the true data generating process while weighting the component models based on their relative performance.

In addition to providing an overall effect estimate, the RoBMA-PSMA model can also be used to calculate Bayes factors (BF) quantifying the strength of support for the presence (or absence) of an intervention effect, heterogeneity, or reporting bias. To illustrate, consider the question of whether there is or is not an intervention effect. Since we used a neutral prior that considered each possibility equally likely, we can take the ratio of the posterior probabilities of model ensembles that included an effect to those that did not. We always report the BFs so they can be interpreted as how many times more likely the data were assuming the best supported hypothesis compared to the other hypothesis. For example, a BF\textsubscript{01} = 2.0 for the absence of an intervention effect suggests that the models assuming all variation is random fit the data twice as well as models predicting an effect is present.\footnote{BF\textsubscript{10} is the inverse of BF\textsubscript{01}, so while BF\textsubscript{01} indicates evidence in support of the null hypothesis, BF\textsubscript{10} indicates evidence in support of the alternative hypothesis.} Bayes factors should be interpreted as a continuous measure of the relative support for one hypothesis versus another, not as the probability a hypothesis is true.

RoBMA-PSMA outperforms other meta-analytic approaches, including each constituent model included in RoBMA-PSMA, in simulations covering a range of plausible scenarios (Maier et al., 2022). Further, an analysis of real data from an approximately known data generating process (multi-lab registered replication report) suggested that RoBMA-PSMA outperforms other available techniques (Bartoš et al., 2022). \textcolor{blue}{Simulations do suggest a simpler version of the RoBMA approach that includes only the PET-PEESE publication bias models (RoBMA-PP) is more effective in scenarios with strong \emph{p}-hacking and publication bias.} These strong performance indices combined with the ability to simultaneously model various plausible manifestations of reporting bias made RoBMA-PSMA \textcolor{blue}{(and RoBMA-PP)} an attractive choice for re-analyzing the enhanced expectancies and self-controlled practice meta-analyses.

@@

The self-controlled practice meta-data,\footnote{\emph{Z}-curve results for self-controlled practice were reported previously in McKay, Yantha, et al. (2022). We reproduce them here for comparison to enhanced expectancies and motivational factors combined analyses.} enhanced expectancy meta-data, and the combination of both the enhanced expectancy and self-controlled practice meta-data were analyzed with a \emph{z}-curve. A \emph{z}-curve analysis estimates the statistical power of all studies ever conducted within a given literature, even if those studies were not reported, on the basis of the significant results that are present (Bartoš \& Schimmack, 2022). That power estimate is equivalent to the expected discovery rate, that is, the expected rate of significant results for a given literature. The expected discovery rate estimated by a \emph{z}-curve and its corresponding 95\% confidence interval can be compared to the observed discovery rate in the literature (the actual rate of significant results). A discrepancy between the 95\% confidence interval of the expected discovery rate estimate and the observed discovery rate provides evidence of reporting bias.

@@

Similar to RoBMA-PSMA (Maier et al., 2022), \emph{z}-curve has also performed well in simulation studies and when applied to multi-lab registered replication data (Bartoš \& Schimmack, 2022). \emph{Z}-curve is a tool that can provide insight into the possible power shortcomings of a particular literature. Further, as \emph{z}-curve was designed to accommodate highly heterogeneous inputs, it is ideal for exploring power and reporting bias among studies that used a variety of methods to manipulate the two motivational factors in Wulf and Lewthwaite's (2016) OPTIMAL theory of motor learning.

@@

We fit two RoBMA-PSMA models to the enhanced expectancies meta-data using the effect sizes and variances calculated by Bacelar and colleagues (2022). The first model included all studies, while the second excluded two influential cases, consistent with the primary results reported in their original meta-analysis. The self-controlled practice effect sizes and standard errors reported by McKay and colleagues (2022) were analyzed using the same strategy.

@@

Statistical analyses were conducted using R (Version 4.2.2; R Core Team, 2021) and the R-packages \emph{geomtextpath} (Version 0.1.0; Cameron \& van den Brand, 2022), \emph{gt} (Version 0.6.0; Iannone et al., 2022), \emph{invgamma} (Version 1.1; Kahle \& Stamey, 2017), \emph{metafor} (Version 3.4.0; Viechtbauer, 2010), \emph{papaja} (Version 0.1.0.9999; Aust \& Barth, 2020), \emph{patchwork} (Version 1.1.0.9000; Pedersen, 2022), \emph{plotly} (Version 4.10.0; Sievert, 2020), \emph{pwr} (Version 1.3.0; Champely, 2020), \emph{renv} (Version 0.15.5; Ushey, 2022), \emph{RoBMA} (Version 2.2.2; Bartoš \& Maier, 2020), \emph{tidyverse} (Version 1.3.1; Wickham et al., 2019), \emph{tinylabels} (Version 0.2.3; Barth, 2022), \emph{truncdist} (Version 1.0.2; Novomestky \& Nadarajah, 2016), and \emph{zcurve} (Version 2.1.2; Bartoš \& Schimmack, 2020) were used in this project.

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure 2B. The results suggest moderate evidence against the presence of an effect, BF\textsubscript{01} = 3.16, very weak evidence against the presence of heterogeneity, BF\textsubscript{rf} = 1.7, and overwhelming evidence for the presence of reporting bias, BF\textsubscript{pb} = 18,399. The overall model ensemble estimated the effect of self-controlled practice as \emph{d} = .034 (95\% credible interval {[}.0, .248{]}). Heterogeneity was estimated as \(\tau\) = .05 (95\% credible interval {[}.0, .261{]}). A model fit with two influential cases\footnote{\textcolor{blue}{The two outliers in the self-controlled practice were studies by Lemos and colleagues (2017) and Marques and colleagues (2017). Lemos et al. 2017 measured ballet movement form and tested the effect of choosing among video demonstrations while Marques et al. 2017 measured front crawl movement form and tested the effect of choosing which self-modeling video to watch during practice.
  }} included found overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 1,924,516 and estimated \(\tau\) = .559 (95\% credible interval {[}.36, .78{]}). There were no other meaningful differences between models.

@@

The model-averaged posterior distribution of the average effect from the RoBMA-PSMA model is displayed in Figure 2C. The results revealed weak evidence for the presence of an effect BF\textsubscript{10} = 1.9, weak evidence for the presence of reporting bias, BF\textsubscript{pb} = 2.3, and very strong evidence for the presence of heterogeneity, BF\textsubscript{rf} = 47.6. The model ensemble estimated an average effect of \emph{d} = .26 (95\% credible interval {[}-.07, .63{]}). Heterogeneity was estimated as \(\tau\) = .35 (95\% credible interval {[}.07, .54{]}). A model fit with two influential cases\footnote{\textcolor{blue}{The two outliers in the enhanced expectancies meta-analysis were Goudini et al. (2018) and Navaee et al. (2018). The study by Goudini and colleagues measured linear tracing performance and tested the effect of feedback after good and bad trials. The study by Navaee and colleagues measured beanbag tossing performance and tested the effect of providing children on the autism spectrum with nonveridical feedback suggesting performance was 20\% better than the average.}} included found weak evidence for the absence of an effect BF\textsubscript{01} = 2.5, strong evidence for the presence of reporting bias, BF\textsubscript{pb} = 21, and overwhelming evidence for the presence of heterogeneity, BF\textsubscript{rf} = 45,300. The model ensemble estimated an average effect of \emph{d} = .00 (95\% credible interval {[}-.62, .59{]}), \(\tau\) = .49 (95\% credible interval {[}.32, .68{]}).

The RoBMA-PSMA models with and without influential cases differed primarily with respect to the fit of the PEESE models. When two large effect sizes with large standard errors were included in the analysis, the PEESE model provided a very strong fit to the data (BF\textsubscript{10} = 39.9) when assuming no true effect but the presence of heterogeneity. When the two influential cases were excluded, the best fitting model was the PET under the same assumptions (BF\textsubscript{10} = 9.4).\footnote{\textcolor{blue}{We also fit RoBMA-PP models that perform better than RoBMA-PSMA models in the presence of strong \emph{p}-hacking and publication bias. The RoBMA-PP models did not lead to substantively different conclusions for the analysis of self-controlled practice or enhanced expectancies.}}

@@

\leavevmode\vadjust pre{\hypertarget{ref-R-RoBMA}{}}%
Bartoš, F., \& Maier, M. (2020). \emph{RoBMA: An r package for robust bayesian meta-analyses}. \url{https://CRAN.R-project.org/package=RoBMA}

@@

\leavevmode\vadjust pre{\hypertarget{ref-lovakov2021}{}}%
Lovakov, A., \& Agadullina, E. R. (2021). Empirically derived guidelines for effect size interpretation in social psychology. \emph{European Journal of Social Psychology}, \emph{00}, 1--20. \url{https://doi.org/10.1002/ejsp.2752}

\leavevmode\vadjust pre{\hypertarget{ref-maier2022}{}}%
Maier, M., Bartoš, F., \& Wagenmakers, E.-J. (2022). Robust {Bayesian} meta-analysis: Addressing publication bias with model-averaging. \emph{Psychological Methods}. \url{https://doi.org/10.1037/met0000405}
